---
title: "`xtune`: Tuning differential shrinkage parameters of penalized regression models based on external information"
author: "Chubing Zeng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview
The main goal of `xtune` is to tuning multiple shrinkage parameters in Lasso and Ridge regression, based on external information. Standard Lasso or Ridge regression only have a single penalty parameter $\lambda$ applied equally to all regression coefficients to control the amount of regularization in the model. Here we apply an individual shrinkage parameter $\lambda_j$ to each regression coefficient $\beta_j$. And the 
vector of shrinkage parameters $\lambda s = (\lambda_1,...,\lambda_p)$ is modeled as a function of the external information $Z$ provided to the model. A better prediction accuracy for penalized regression models may be achieved by allowing individual shrinkage for each regression coefficients based on external in-formation. 

We employ an Empirical Bayes approach to learn the multiple tuning parameters. Once the tuning parameters $\lambda$s are estimated, and therefore the penalties known, the regression coefficients are obtained using glmnet.

The response variable can be quantitative a factor with two levels. For binary response variable, a sparse LDA procedure is used to get the fisher discriminant direction and the discriminant function for a new data point. Utilities for carrying out post-fitting visualization, summarization, and prediction are also provided.


## Examples

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
